Fourmile Calibrations and Output

### 2025-02-25 fourmile_output_dec2024.Rmd
updated water partition boxplots. need to finish adding tukey letters. added 2015 as example year for cumulative flows

@@

gw_maxcumsum <- gw %>%
  group_by(source, year) %>%
  summarize(yr_cum_RD1 = max(RD1cumsum),
            yr_cum_RD2 = max(RD2cumsum),
            yr_cum_RG1 = max(RG1cumsum),
            yr_cum_RG2 = max(RG2cumsum)) %>%
  mutate(year = as.factor(year)) %>%
  pivot_longer(!c(year, source), names_to = 'flow_type', values_to = 'flow_cms')

ggplot(gw_maxcumsum, aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1)) +
  #geom_point(position=position_jitterdodge(),alpha=0.3) +
  ggtitle('Annual sum') +
  labs(y = 'mm of water') + theme_bw(base_size=20)

ggplot(gw_maxcumsum, aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1)) +
  #geom_point(position=position_jitterdodge(),alpha=0.3) +
  #geom_point(aes(color = source)) +
  facet_wrap(~flow_type, scales = 'free') +
  ggtitle('Annual sum') +
  labs(y = 'mm of water') + theme_bw(base_size=20)

>>>>

# look at 2015, high snow year 
example <- gw %>%
  filter(year == 2015) %>%
  ungroup() %>% 
  dplyr::select(date, source,
         RD1cumsum,
         RD2cumsum,
         RG1cumsum,
         RG2cumsum) %>%
  pivot_longer(!c(date,source), names_to='flow_type',
               values_to='flow_cms') %>%
  mutate(source = if_else(source == 'grid_all',
                          'Gridded',
                          'Station'))

facet_labels <- c('RD1cumsum' = 'Surface runoff',
                  'RD2cumsum' = 'Interflow',
                  'RG1cumsum' = 'Fast groundwater',
                  'RG2cumsum' = 'Slow groundwater')

ggplot(example,
       aes(x=date, y = flow_cms, color=source)) +
  geom_line() +
  facet_wrap(~flow_type, 
             scales = 'free_y',
             labeller = labeller(flow_type = facet_labels)) +
   geom_line(size=1, alpha=0.7) +  
  scale_color_manual(values = c("#0047AB", "orange")) +
  labs(x=element_blank(),
       y='Flow (cms)') +
  theme_bw(base_size = 20) +
  theme(
    legend.position = c(0.94, 0.08),
    legend.text = element_text(size = 12),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black")
  )

ggsave('./figures/2015_flowcumulative.png',
       dpi=600)

@@

gw_maxcumsum <- gw %>%
  group_by(source, year) %>%
  summarize(yr_cum_RD1 = max(RD1cumsum),
            yr_cum_RD2 = max(RD2cumsum),
            yr_cum_RG1 = max(RG1cumsum),
            yr_cum_RG2 = max(RG2cumsum)) %>%
  mutate(year = as.factor(year)) %>%
  pivot_longer(!c(year, source), names_to = 'flow_type', values_to = 'flow_cms') %>%
  mutate(source = if_else(source == 'grid_all',
                          'Gridded',
                          'Station'))

ggplot(gw_maxcumsum, aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1)) +
  #geom_point(position=position_jitterdodge(),alpha=0.3) +
  scale_y_log10() +
  ggtitle('Annual sum') +
  labs(y = 'mm of water') + theme_bw(base_size=20) # this plot doesnt show lower values (eg rd1)
# i tried with log scale on y and it looks really bad so going with facet below

facet_labels <- c('yr_cum_RD1' = 'Surface runoff',
                  'yr_cum_RD2' = 'Interflow',
                  'yr_cum_RG1' = 'Fast groundwater',
                  'yr_cum_RG2' = 'Slow groundwater')

ggplot(gw_maxcumsum, aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1),
               outliers=F) + # need to note 1 outlier is removed from RD1 grid and statn (2013 flood) ****
  stat_compare_means(data = gw_maxcumsum %>% 
                       filter(flow_type != "yr_cum_RD1"),
                     aes(group = source),
                     method = "wilcox.test",
                     label = "p.signif",
                     hide.ns = T,
                     paired = F,
                     size = 7) +
  facet_wrap(~flow_type, 
             scales = 'free',
             labeller = labeller(flow_type = facet_labels)) +
  labs(y = 'mm of water') + 
  theme_bw(base_size=20) +
  scale_fill_manual(values = c("#0047AB", "orange")) +
  theme(
    legend.position = c(0.9, 0.38),
    legend.text = element_text(size = 12),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.15)))

ggsave('./figures/water_partition_boxplots.png',
       dpi=600)

# want to see distribution
ggplot(gw_maxcumsum, aes(x = flow_cms, fill = source)) +
  geom_density(alpha = 0.6, adjust = 1.2) +  # Adjust smoothness with `adjust`
  facet_wrap(~flow_type, 
             scales = 'free', 
             labeller = labeller(flow_type = facet_labels)) +
  labs(x = "Flow (cms)", y = "Density") + 
  theme_bw(base_size = 20) +
  scale_fill_manual(values = c("#0047AB", "orange")) +
  theme(
    legend.position = 'top',
    legend.text = element_text(size = 16),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black")
  )

### RD and RG stats

model <-aov(flow_cms ~ source * flow_type, 
            data = gw_maxcumsum)

# Tukey HSD
tukey_test <- TukeyHSD(model)

tukey_test

library(multcompView)
tukey_letters <- multcompLetters4(model, tukey_test)

tukey_letters$`source:flow_type`$Letters

# make a vector of the letters
letters_vector <- c("a", "b", "b", "c", "c", "c", "c", "c")

gw_maxcumsum <- gw_maxcumsum %>%
  mutate(letter = case_when(
    source == "Gridded" & flow_type == "yr_cum_RG1" ~ letters_vector[1],
    source == "Station" & flow_type == "yr_cum_RG1" ~ letters_vector[2],
    source == "Station" & flow_type == "yr_cum_RD2" ~ letters_vector[3],
    source == "Gridded" & flow_type == "yr_cum_RD1" ~ letters_vector[4],
    source == "Station" & flow_type == "yr_cum_RD1" ~ letters_vector[5],
    source == "Gridded" & flow_type == "yr_cum_RD2" ~ letters_vector[6],
    source == "Gridded" & flow_type == "yr_cum_RG2" ~ letters_vector[7],
    source == "Station" & flow_type == "yr_cum_RG2" ~ letters_vector[8]
  ))

# now make the same figure as above but with letters
facet_labels <- c('yr_cum_RD1' = 'Surface runoff',
                  'yr_cum_RD2' = 'Interflow',
                  'yr_cum_RG1' = 'Fast groundwater',
                  'yr_cum_RG2' = 'Slow groundwater')

gw_maxcumsum_letters <- gw_maxcumsum %>%
  group_by(source, flow_type) %>%
  mutate(max = max(flow_cms)) %>%
  distinct(flow_type, source, .keep_all = TRUE) %>%
    mutate(max = if_else((flow_type == 'yr_cum_RD1' & 
                           source == 'Gridded'),
                         3.5, max)) %>%
    mutate(max = if_else((flow_type == 'yr_cum_RD1' & 
                           source == 'Station'),
                         1.3, max))

ggplot(gw_maxcumsum, aes(x = flow_type, y = flow_cms, fill = source)) +
  geom_boxplot(position = position_dodge(1),
               outliers=F) + # need to note 1 outlier is removed from RD1 grid and statn (2013 flood) ****
  geom_text(data = gw_maxcumsum_letters, 
            aes(label=letter,
                x=flow_type, 
                y=max), 
            position=position_dodge2(0.75), 
            vjust = -6) + 
  facet_wrap(~flow_type, 
             scales = 'free',
             labeller = labeller(flow_type = facet_labels)) +
  labs(y = 'mm of water') + 
  theme_bw(base_size=20) +
  scale_fill_manual(values = c("#0047AB", "orange")) +
  theme(
    legend.position = c(0.9, 0.38),
    legend.text = element_text(size = 12),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank())

ggsave('./figures/water_partition_boxplots.png',
       dpi=600)
